Compositional Verification of Concurrent C Programs with Search Structure Templates

### 2023-05-02 concurrency/CONCUR 2023/paper.tex
move insert to section templates

The definitions of $\nodeboxrep$ and $\treerep$ may vary depending on the data structure, but fundamentally, $\nodeboxrep$ holds the resources owned by each client thread, while $\treerep$ holds the abstract representation of the data structure, treated as an atomically accessed shared resource. In particular, holding $\nodeboxrep(\texttt{n})$ should allow us to acquire the lock on node \texttt{n}, which in turn gives us access to the contents of the node. Formally, we write
\[\nodeboxrep(\texttt{n}) \triangleq\ \texttt{n->lock} \lockvar \mathsf{R}(\texttt{n})\]
where the \emph{lock invariant assertion} $\ell \lockvar \mathsf{R}$ expresses that the lock $\ell$ protects resources $\mathsf{R}$. In this case, $\mathsf{R}$ contains the contents of the node (i.e., the \lstinline{node} struct contained inside the \lstinline{node_t} struct of \lstinline{pn->n}).

@@

@@

@@

\subsection{Lock-coupling BST}
\label{BST_lock_insert}
\begin{figure}[h]
	\begin{subfigure}[t]{0.48\textwidth}
	\lstinputlisting[language=C, style=myStyle, escapechar=|]{lock_insert.c} 
	\caption{The \lstinline{insert} method of the lock-coupling template algorithm}
	\label{insert_lock}	
	\end{subfigure}\qquad
    \begin{subfigure}[t]{0.48\textwidth}
    \lstinputlisting[language=C, style=myStyle]{giveup_insert.c} 
    \caption{The \lstinline{insert} method of the give-up template algorithm}
    \label{insert_giveup}
    \end{subfigure} 
\caption{The \lstinline{insert} method of the lock-coupling and give-up template algorithms}
\label{insert_lock_giveup} 
\end{figure}

Figure \ref{insert_lock} shows the implementation of \lstinline{insert} for the lock-coupling BST. \wm{Actually, is this \lstinline{insert} function part of the template? If so, should we discuss it in section 3?} It makes use of the \texttt{traverse} function of Section \ref{lock-coupling-algo} to find the node at which to insert the key \lstinline{k}. If \texttt{traverse} returns 0, it has reached a node containing key \texttt{k}, and we only have to change the node's value to the target \texttt{value} (lines 5-6 in Figure \ref{insert_lock}). Otherwise, \texttt{traverse} has reached an empty leaf that can hold key \texttt{k}, so it calls the data-structure-specific \texttt{insertOp} function to allocate a new node with key \texttt{k} and value \texttt{v} (line 9 in Figure \ref{insert_lock}).

\subsection{Give-up BST}
\label{BST_giveup_insert}
The give-up version of the \texttt{insert} method (see Figure \ref{insert_giveup}) is almost identical to the lock-coupling version. The one difference is that it does not require acquiring a lock before calling its \texttt{traverse} function.

The give-up BST uses the same definition and specification of \texttt{findNext}, while introducing an extra helper function called \texttt{inRange} \wm{If \lstinline{inRange} is part of the template, I think we should discuss it in the previous section}. \than{If we move it, this section is very short.} The \texttt{insertOp} function is slightly different from that of the lock-coupling template, as each \lstinline{node_t} struct has two additional fields, \texttt{min} and \texttt{max}, to implement \lstinline{inRange}, as described in Section \ref{give-up-algo}.

@@

@@

The proof of the \lstinline{insert} function is similar to that of the previous function: the only difference is that the give-up version of \lstinline{insert} does not acquire a lock before calling \lstinline{traverse}, reflecting the difference in the locking patterns and the preconditions of \lstinline{traverse} between the two templates. In Appendix \ref{sec:apd_proof}, Figure \ref{proof_giveup_insert} presents an annotation that uses separation logic for the \texttt{insert} function within the give-up template for BST.

>>>>

The definitions of $\nodeboxrep$ and $\treerep$ may vary depending on the data structure, but fundamentally, $\nodeboxrep$ holds the resources owned by each client thread, while $\treerep$ holds the abstract representation of the data structure, treated as an atomically accessed shared resource. In particular, holding $\nodeboxrep(\texttt{n})$ should allow us to acquire the lock on node \texttt{n}, which in turn gives us access to the contents of the node. Formally, we write
\[\nodeboxrep(\texttt{n}) \triangleq\ \texttt{n->lock} \lockvar \mathsf{R}(\texttt{n})\]
where the \emph{lock invariant assertion} $\ell \lockvar \mathsf{R}$ expresses that the lock $\ell$ protects resources $\mathsf{R}$. In this case, $\mathsf{R}$ contains the contents of the node (i.e., the \lstinline{node} struct contained inside the \lstinline{node_t} struct of \lstinline{n}).

@@

\than{I moved \texttt{insert} from Section 4 to here. William: Could you please write something here to have a smooth transition from \texttt{traverse} to \texttt{insert}, or do you think you need to have subsections for \texttt{traverse} and \texttt{insert}? }

\begin{figure}[h]
	\begin{subfigure}[t]{0.48\textwidth}
		\lstinputlisting[language=C, style=myStyle, escapechar=|]{lock_insert.c} 
		\caption{The \lstinline{insert} method of the lock-coupling template algorithm}
		\label{insert_lock}	
	\end{subfigure}\qquad
	\begin{subfigure}[t]{0.48\textwidth}
		\lstinputlisting[language=C, style=myStyle]{giveup_insert.c} 
		\caption{The \lstinline{insert} method of the give-up template algorithm}
		\label{insert_giveup}
	\end{subfigure} 
	\caption{The \lstinline{insert} method of the lock-coupling and give-up template algorithms}
	\label{insert_lock_giveup} 
\end{figure}

\than{Should we move the code to template section?}

Figure \ref{insert_lock} shows the implementation of \lstinline{insert} for the lock-coupling template. %\wm{Actually, is this \lstinline{insert} function part of the template? If so, should we discuss it in section 3?} 
It makes use of the \texttt{traverse} function %of Section \ref{lock-coupling-algo} 
to find the node at which to insert the key \lstinline{k}. If \texttt{traverse} returns 0, it has reached a node containing key \texttt{k}, and we only have to change the node's value to the target \texttt{value} (lines 5-6 in Figure \ref{insert_lock}). Otherwise, \texttt{traverse} has reached an empty leaf that can hold key \texttt{k}, so it calls the data-structure-specific \texttt{insertOp} function to allocate a new node with key \texttt{k} and value \texttt{v} (line 9 in Figure \ref{insert_lock}).

@@

\than{I moved \texttt{insert} from Section 4 to here. William: Could you please write something here to have a smooth transition from \texttt{traverse} to \texttt{insert}, or do you think you need to have subsections for \texttt{traverse} and \texttt{insert}? }

The give-up version of the \texttt{insert} method (see Figure \ref{insert_giveup}) is almost identical to the lock-coupling version. The one difference is that it does not require acquiring a lock before calling its \texttt{traverse} function. The proof of the \lstinline{insert} function is similar to that of the previous function: the only difference is that the give-up version of \lstinline{insert} does not acquire a lock before calling \lstinline{traverse}, reflecting the difference in the locking patterns and the preconditions of \lstinline{traverse} between the two templates. In Appendix \ref{sec:apd_proof}, Figure \ref{proof_giveup_insert} presents an annotation that uses separation logic for the \texttt{insert} function within the give-up template for BST.

@@

The give-up BST uses the same definition and specification of \texttt{findNext}, while introducing an extra helper function called \texttt{inRange} \wm{If \lstinline{inRange} is part of the template, I think we should discuss it in the previous section}. The \texttt{insertOp} function is slightly different from that of the lock-coupling template, as each \lstinline{node_t} struct has two additional fields, \texttt{min} and \texttt{max}, to implement \lstinline{inRange}, as described in Section \ref{give-up-algo}.

@@

\subsection{Lock-coupling BST}
\label{BST_lock_insert}
\than{What do we need to say here? (1) - still keep subsection and filled \texttt{insertOp} and the way we prove \texttt{insert} here, or (2) - get rid of this subsection.}
\subsection{Give-up BST}
\label{BST_giveup_insert}
\than{What do we need to say here? (1) - still keep subsection and filled \texttt{insertOp}, and \texttt{inRange} and the way we prove \texttt{insert} here, or (2) - get rid of this subsection.}

@@
